A Gaussian process model of human e\-lect\-ro\-cort\-i\-co\-graph\-ic data

### 2020-03-09 paper/main.tex
next round of JRM edits

\title{A Gaussian process model of human electrocorticographic data}
\author{
  Lucy L. W. Owen$^{1}$,
  Tudor A. Muntianu$^{1}$,
  Andrew C. Heusser$^{1, 2}$, \\
  Patrick Daly$^{3}$,
  Katherine Scangos$^{3}$, and
  Jeremy R. Manning$^{1\ast}$\\\\
$^{1}$Department of Psychological and Brain Sciences, Dartmouth College,\\
Hanover, NH 03755, USA\\
$^{2}$Akili Interactive,\\
Boston, MA 02110, USA\\
$^{3}$Weill Institute for Neurosciences, University of California, San Francisco,\\
San Francisco, CA 94121, USA}

@@

An increasingly popular approach is to improve the effective spatial
resolution of MEG or scalp EEG data by using a geometric approach
called \textit{beamforming} to solve the biomagnetic or bioelectrical
inverse problem~\citep{Sarv87}.  This approach entails using detailed
brain conductance models (often informed by high spatial resolution
anatomical MRI images) along with the known sensor placements
(localized precisely in 3D space) to reconstruct brain signals
originating from theoretical point sources deep in the brain (and far
from the sensors).  Traditional beamforming approaches must overcome
two obstacles.  First, the inverse problem beamforming seeks to solve
has infinitely many solutions.  Researchers have made traction towards
constraining the solution space by assuming that signal-generating
sources are localized on a regularly spaced grid spanning the brain
and that individual sources are small relative to their distances to
the sensors~\citep{Snyd91, BailEtal01, HillEtal05}.  The second, and in
some ways much more serious, obstacle is that the magnetic fields
produced by external (noise) sources are substantially stronger than
those produced by the neuronal changes being sought (i.e., at deep
structures, as measured by sensors at the scalp).  This means that
obtaining adequate signal quality often requires averaging the
measured responses over tens to hundreds of responses or trials~\citep[e.g., see review by][]{HillEtal05}.

@@

The current ``gold standard'' for precisely localizing signals and sampling at
high temporal resolution is to take (ECoG or iEEG) recordings from implanted
electrodes (but from a limited set of locations in any given brain).  This begs
the following question: what can we infer about the activity exhibited by the
rest of a person's brain, given what we learn from the limited intracranial
recordings we have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we call
\textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by Robert J. Sawyer
in his popular science fiction novel \textit{The Terminal
Experiment}~\cite{Sawy95}.  SuperEEG is a fictional technology that measures
ongoing neural activity throughout the entire living human brain with perfect
precision and at arbitrarily high spatiotemporal resolution.}, based on Gaussian
process regression~\citep{Rasm06}.  SuperEEG entails using data from multiple
people to estimate activity patterns at arbitrary locations in each person's
brain (i.e., independent of their electrode placements).  We test our SuperEEG
approach using two large datasets of intracranial recordings~\citep{SedeEtal03,
SedeEtal07a, SedeEtal07b, MannEtal11, MannEtal12, EzzyEtal17, HoraEtal17,
KragEtal17, KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18,
KuceEtal18}.  We show that the SuperEEG algorithm recovers signals well from
electrodes that were held out of the training dataset.  We also examine the
factors that influence how accurately activity may be estimated (recovered),
which may have implications for electrode design and placement in neurosurgical
applications.

\section*{Approach} The SuperEEG approach to inferring high temporal resolution
full-brain activity patterns is outlined and summarized in
Figure~\ref{fig:methods}. We describe (in this section) and evaluate (in
\textit{Results}) our approach using a two large previously collected datasets
comprising multi-session intracranial recordings. Dataset 1 comprises
multi-session recordings taken from 6876 electrodes implanted in the brains of
88 epilepsy patients~\citep{SedeEtal03, SedeEtal07a, SedeEtal07b, MannEtal11,
MannEtal12}.  Each recording session lasted from 0.2--3~h (total recording time:
0.3--14.2~h; Fig.~\suppstats E).  During each recording session, the patients
participated in a free recall list learning task, which lasted for up to
approximately 1~h.  In addition, the recordings included ``buffer'' time (the
length varied by patient) before and after each experimental session, during
which the patients went about their regular hospital activities (confined to
their hospital room, and primarily in bed).  These additional activities
included interactions with medical staff and family, watching television,
reading, and other similar activities.  For the purposes of the Dataset 1
analyses presented here, we aggregated all data across each recording session,
including recordings taken during the main experimental task as well as during
non-experimental time.  We used Dataset 1 to develop our main SuperEEG approach,
and to examine the extent to which SuperEEG might be able to generate
task-general predictions.  Dataset 2 comprised multi-session recordings from
14860 electrodes implanted in the brains of 131 epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17, LinEtal17,
SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Each recording session lasted
from 0.4--2.2~h (total recording time: 0.4--6.6~h; Fig.~\suppstats K).  Whereas
Dataset 1 included recordings taken as the patients participated in a variety of
activities, Dataset 2 included recordings taken as each patient performed each
of two specific experimental memory tasks: a random word list free recall task
(Experiment 1) and a categorized word list free recall task (Experiment 2).  We
used Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides additional
information about both datasets.

\begin{figure}
  \centering \includegraphics[width=\textwidth]{figs/methods}
  \caption{\textbf{Methods overview.}  \textbf{A.~Electrode locations.}  Each
  dot reflects the location of a single electrode implanted in the brain of a
  Dataset 1 patient.  A held-out recording location from one patient is
  indicated in red, and the patient's remaining electrodes are indicated in
  black. The electrodes from the remaining patients are colored by $k$-means
  cluster (computed using the full-brain correlation model shown in Panel D).
  \textbf{B.~Radial basis function kernel.} Each electrode contributed by the
  patient (black) weights on the full set of locations under consideration (all
  dots in Panel A, defined as $\bar{R}$ in the text).  The weights fall off with
  positional distance (in MNI space) according to an RBF. \textbf{C.~Per-patient
  correlation matrices.}  After computing the pairwise correlations between the
  recordings from each patient's electrodes, we use RBF-weighted averages to
  estimate correlations between all locations in $\bar{R}$.  We obtain an
  estimated full-brain correlation matrix using each patient's data.
  \textbf{D.~Merged correlation model.}  We combine the per-patient correlation
  matrices (Panel C) to obtain a single full-brain correlation model that
  captures information contributed by every patient.  Here we have sorted the
  rows and columns to reflect $k$-means clustering labels~\citep[using
  $k$=7;][]{YeoEtal11}, whereby we grouped locations based on their correlations
  with the rest of the brain (i.e., rows of the matrix displayed in the panel).
  The boundaries denote the cluster groups.  The rows and columns of Panel C
  have been sorted using the Panel D-derived cluster labels.
  \textbf{E.~Reconstructing activity throughout the brain.}  Given the observed
  recordings from the given patient (shown in black; held-out recording is shown
  in blue), along with a full-brain correlation model (Panel D), we use
  Equation~\ref{eqn:reconstruction} to reconstruct the most probable activity at
  the held-out location (red).} \label{fig:methods}
\end{figure}

We first applied fourth order Butterworth notch filters to remove 60 Hz ($\pm$
.5 Hz) line noise from every recording (from every electrode). Next, we
downsampled the recordings (regardless of the original sample rate) to 250 Hz.
This downsampling step served to both normalize for differences in sampling
rates across patients and to ease the computational burden of our subsequent
analyses.  We then excluded any electrodes that showed putative epileptiform
activity. Specifically, we excluded from further analysis any electrode that
exhibited a maximum kurtosis of 10 or greater across all of that patient's
recording sessions.  We also excluded any patients with fewer than 2 electrodes
that passed this criteria, as the SuperEEG algorithm requires measuring
correlations between 2 or more electrodes from each patient.  For Dataset 1,
this yielded clean recordings from 4168 electrodes implanted throughout the
brains of 67 patients (Fig.~\ref{fig:methods}A, colored dots); for Dataset 2,
this yielded clean recordings from 5023 electrodes implanted in 78 patients.
Each individual patient contributed electrodes from a limited set of brain
locations, which we localized in a common space~\citep[MNI152;][]{GrabEtal06};
an example Dataset 1 patient's 54 electrodes that survived the kurtosis
thresholding procedure are highlighted in black and red
(Fig.~\ref{fig:methods}A).

@@

We designed our approach to be agnostic to electrode impedances, as electrodes
that do not exist do not have impedances.  Therefore our algorithm recovers
voltages in standard deviation ($z$-scored) units rather than attempting to
recover absolute voltages. (This property reflects the fact that
$\hat{K}_{\beta_s, \alpha_s}$ and $\hat{K}_{\alpha_s, \alpha_s}$ are correlation
matrices rather than covariance matrices.)  Also, we note that
Equation~\ref{eqn:reconstruction} requires computing a $T$ by $T$ matrix, which
can become computationally intractable when $T$ is very large (e.g., an example
patient in Dataset 1, $T = 51147$; also see Fig.~\suppstats, Panels E and
K). However, because Equation~\ref{eqn:reconstruction} is time invariant, we may
compute $Y_{s,k,\beta_s}$ in a piecewise manner by filling in $Y_{s,k,\beta_s}$
one row at a time (using the corresponding samples from $Y_{s, k, \alpha_s}$).

@@

We used this general approach to quantify the algorithm's performance across the
full dataset. For each held-out electrode, from each held-out patient in turn,
we computed the average correlation (across recording sessions) between the
SuperEEG-reconstructed voltage traces and the observed voltage traces from that
electrode.  For this analysis we set $\bar{R}$ to be the union of all electrode
locations across all patients.  This yielded a single correlation coefficient
for each electrode location in $\bar{R}$, reflecting how well the SuperEEG
algorithm was able to recover the recording at that location by incorporating
data across patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations was
centered well above zero (mean: 0.51; $t$-test comparing mean of distribution of
$z$-transformed average patient correlation coefficients to 0: $t(66) = 23.55, p
< 10^{-10}$), indicating that the SuperEEG algorithm recovers held-out activity
patterns substantially better than random guessing.

Next, we compared the quality of these across-participant reconstructions (i.e.,
computed using a correlation model learned from other patients' data) to
reconstructions generated using a correlation model trained using the
in-patient's data.  In other words, for this within-patient benchmark analysis
we estimated $\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in
turn, using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions serve as
an estimate of how well data from all of the other electrodes from that single
patient may be used to estimate held-out data from the same patient.  This
allows us to ask how much information about the activity at a given electrode
might be inferred through (a) volume conductance or other sources of ``leakage''
from activity patterns measured from the patient's other electrodes and (b)
across-electrode correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of within-patient
correlations was centered well above zero (mean: 0.32; $t$-test comparing mean
of distribution of $z$-transformed average patient correlation coefficients to
0: $t(66) = 15.16, p < 10^{-10}$). However, the across-patient correlations were
substantially higher ($t$-test comparing average $z$-transformed within versus
across patient electrode correlations: $t(66) = 9.17, p < 10^{-10}$).  This is
an especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all data from
the patient whose data is being reconstructed.  We repeated each of these
analyses on a second independent dataset and found similar results
(Fig.~\ref{fig:corrmap}B, D; within versus across reconstruction accuracy:
$t(77) = 11.25, p < 10^{-10}$). We also replicated this result separately for
each of the two experiments from Dataset 2 (Fig.~\perexpcorrmaps).  This overall
finding, that reconstructions of held-out data using correlation models learned
from \textit{other} patient's data yield higher reconstruction accuracy than
correlation models learned from the patient whose data is being reconstructed,
has two important implications.  First, it implies that distant electrodes
provide additional predictive power to the data reconstructions beyond the
information contained solely in nearby electrodes.  (This follows from the fact
that each patient's grid, strip, and depth electrodes are implanted in a unique
set of locations, so for any given electrode the closest electrodes in the full
dataset tend to come from the same patient.)  Second, it implies that the
spatial correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

The recordings we analyzed from Dataset 1 comprised data collected as the
patients performed a variety of (largely idiosyncratic) tasks throughout each
day's recording session.  That we observed reliable reconstruction accuracy
across patients suggests that the spatial correlations derived from the SuperEEG
algorithm are, to some extent, similar across tasks.  We tested this finding
more directly using Dataset 2.  In Dataset 2, the recordings were limited to
times when each patient was participating in one of two experiments.  Experiment
1 is a random-word list free recall task; Experiment 2 is a categorized list
free recall task (24 patients particpated in both).  We wondered whether a
correlation model learned from data from one experiment might yield good
predictions of data from the other experiment.  Further, we wondered about the
extent to which it might be beneficial or harmful to combine data across tasks.

To test the task-specificity of the SuperEEG-derived correlation models, we
restricted the dataset to the 24 patients that participated in both experiments
and repeated the above within- and across-patient cross validation procedures
separately for Experiment 1 and Experiment 2 data from Dataset 2.  We then
compared the reconstruction accuracies for held-out electrodes, for models
trained within versus across the two experiments, or combining across both
experiments (Fig.~\perexptaskrecon).  In every case we found that across-patient
models trained using data from all other patients out-performed within-patient
models trained on data only from the subject contributing the given electrode
($t$s$(23) > 6.50, p$s$ < 10^{-5}$). All reconstruction accuracies also reliably
exceeded chance performance ($t$s$(23) > 8.00, p$s$ < 10^{-8}$).  Average
reconstruction accuracy was highest for the across-patient models limited to
data from the same experiment (mean accuracy: $0.68$); next-highest for the
models that combined data across both experiments (mean accuracy: $0.61$); and
lowest for models trained across tasks (mean accuracy: $0.47$).  This pattern of
results also held for each of the Dataset 2 experiments individually
(Fig.~\perexptaskreconseparated).  Taken together, these results indicate that
there are reliable commonalities in the spatial correlations of full-brain
activity across tasks, but that there are also reliable differences in these
spatial correlations across tasks. Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is highest when
constrained to within-task data, or data that includes a variety of tasks (e.g.,
Dataset 1, or combining across the two Dataset 2 experiments).

Although both datasets we examined provide good full-brain coverage (when
considering data from every patient), electrodes were not sampled uniformly
throughout the brain.  For example, electrodes are more likely to be implanted
in regions like the medial temporal lobe (MTL), and are rarely implanted in
occipital cortex (Fig.~\ref{fig:density}A, B). Separately for each dataset, for
each voxel in the 4~mm$^3$ voxel MNI152 brain, we computed the proportion of
electrodes in the dataset that were contained within a 20 MNI unit radius sphere
centered on that voxel.  We defined the \textit{density} at that location as
this proportion. Across Datasets 1 and 2, the electrode placement densities were
similar (correlation by voxel: $r = 0.6, p < 10^{-10}$).  We wondered whether
regions with good covererage might be associated with better reconstruction
accuracy. For example, Figures~\ref{fig:density}A and B indicate that many
electrodes in the MTL (which tends to be relatively densely sampled) have
relatively high reconstruction accuracy, and occipital electrodes (which tends
to be relatively sparsely sampled) tend to have relatively low reconstruction
accuracy.  To test whether this held more generally across the entire brain, for
each dataset we computed the electrode placement density for each electrode from
each patient (using the proportion of \textit{other} patients' electrodes within
20 MNI units of the given electrode).  We then correlated these density values
with the across-patient reconstruction accuracies for each electrode.  We found
no reliable correlation between reconstruction accuracy and density for Dataset
1 ($r = 0.05, p = 0.70$) and a reliable negative correlation for Dataset 2 ($r =
-0.21, p = 0.05$).  This indicates that the reconstruction accuracies we
observed are \textit{not} driven solely by sampling density, but rather may also
reflect higher order properties of neural dynamics such as functional
correlations between distant voxels~\citep{BetzEtal17b}.

>>>>

\title{A Gaussian process model of human e\-lect\-ro\-cort\-i\-co\-graph\-ic data}
\author{
  Lucy L. W. Owen$^{1}$,
  Tudor A. Muntianu$^{1}$,
  Andrew C. Heusser$^{1, 2}$, \\
  Patrick Daly$^{3}$,
  Katherine Scangos$^{3}$, and
  Jeremy R. Manning$^{1\ast}$\\\\
$^{1}$Department of Psychological and Brain Sciences, Dartmouth College,\\
Hanover, NH 03755, USA\\
$^{2}$Akili Interactive,\\
Boston, MA 02110, USA\\
$^{3}$Weill Institute for Neurosciences, University of California, San Francisco,\\
San Francisco, CA 94121, USA}

@@

An increasingly popular approach is to improve the effective spatial
resolution of MEG or scalp EEG data by using a geometric approach
called \textit{beamforming} to solve the biomagnetic or bioelectrical
inverse problem~\citep{Sarv87}.  This approach entails using detailed
brain conductance models (often informed by high spatial resolution
anatomical MRI images) along with the known sensor placements
(localized precisely in 3D space) to reconstruct brain signals
originating from theoretical point sources deep in the brain (and far
from the sensors).  Traditional beamforming approaches must overcome
two obstacles.  First, the inverse problem beamforming seeks to solve
has infinitely many solutions.  Researchers have made progress towards
constraining the solution space by assuming that signal-generating
sources are localized on a regularly spaced grid spanning the brain
and that individual sources are small relative to their distances to
the sensors~\citep{Snyd91, BailEtal01, HillEtal05}.  The second, and in
some ways much more serious, obstacle is that the magnetic fields
produced by external (noise) sources are substantially stronger than
those produced by the neuronal changes being sought (i.e., at deep
structures, as measured by sensors at the scalp).  This means that
obtaining adequate signal quality often requires averaging the
measured responses over tens to hundreds of responses or trials~\citep[e.g., see review by][]{HillEtal05}.

@@

The current ``gold standard'' for precisely localizing signals and sampling at
high temporal resolution is to take (ECoG or iEEG) recordings from implanted
electrodes (but from a limited set of locations in any given brain).  This begs
the following question: what can we infer about the activity exhibited by the
rest of a person's brain, given what we learn from the limited intracranial
recordings we have from their brain and additional recordings taken from
\textit{other} people's brains?  Here we develop an approach, which we call
\textit{SuperEEG}\footnote{The term ``SuperEEG'' was coined by Robert J. Sawyer
in his popular science fiction novel \textit{The Terminal
Experiment}~\citep{Sawy95}.  SuperEEG is a fictional technology that measures
ongoing neural activity throughout the entire living human brain at arbitrarily
high spatiotemporal resolution.}, based on Gaussian process
regression~\citep{Rasm06}.  SuperEEG entails using data from multiple people to
estimate activity patterns at arbitrary locations in each person's brain (i.e.,
independent of their electrode placements).  We test our SuperEEG approach using
two large datasets of intracranial recordings~\citep{SedeEtal03, SedeEtal07a,
SedeEtal07b, MannEtal11, MannEtal12, EzzyEtal17, HoraEtal17, KragEtal17,
KuceEtal17, LinEtal17, SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  We show
that the SuperEEG algorithm recovers signals well from electrodes that were held
out of the training dataset.  We also examine the factors that influence how
accurately activity may be estimated (recovered), which may have implications
for electrode design and placement in neurosurgical applications.

\section*{Approach}

The SuperEEG approach to inferring high temporal resolution full-brain activity
patterns is outlined and summarized in Figure~\ref{fig:methods}. We describe (in
this section) and evaluate (in \textit{Results}) our approach using two large
previously collected datasets comprising multi-session intracranial recordings.
Dataset 1 comprises multi-session recordings taken from 6876 electrodes
implanted in the brains of 88 epilepsy patients~\citep{SedeEtal03, SedeEtal07a,
SedeEtal07b, MannEtal11, MannEtal12}.  Each recording session lasted from
0.2--3~h (total recording time: 0.3--14.2~h; Fig.~\suppstats E).  During each
recording session, the patients participated in a free recall list learning
task, which lasted for up to approximately 1~h.  In addition, the recordings
included ``buffer'' time (the length varied by patient) before and after each
experimental session, during which the patients went about their regular
hospital activities (confined to their hospital room, and primarily in bed).
These additional activities included interactions with medical staff and family,
watching television, reading, and other similar activities.  For the purposes of
the Dataset 1 analyses presented here, we aggregated all data across each
recording session, including recordings taken during the main experimental task
as well as during non-experimental time.  We used Dataset 1 to develop our main
SuperEEG approach, and to examine the extent to which SuperEEG might be able to
generate task-general predictions.  Dataset 2 comprised multi-session recordings
from 14860 electrodes implanted in the brains of 131 epilepsy
patients~\citep{EzzyEtal17, HoraEtal17, KragEtal17, KuceEtal17, LinEtal17,
SoloEtal18, WeidEtal18, EzzyEtal18, KuceEtal18}.  Each recording session lasted
from 0.4--2.2~h (total recording time: 0.4--6.6~h; Fig.~\suppstats K).  Whereas
Dataset 1 included recordings taken as the patients participated in a variety of
activities, Dataset 2 included recordings taken as each patient performed each
of two specific experimental memory tasks: a random word list free recall task
(Experiment 1) and a categorized word list free recall task (Experiment 2).  We
used Dataset 2 to further examine the ability of SuperEEG to generalize its
predictions within versus across tasks.  Figure~\suppstats~provides additional
information about both datasets.

\begin{figure}
  \centering \includegraphics[width=\textwidth]{figs/methods}
  \caption{\textbf{Methods overview.}  \textbf{A.~Electrode locations.}  Each
  dot reflects the location of a single electrode implanted in the brain of a
  Dataset 1 patient.  A held-out recording location from one patient is
  indicated in red, and the patient's remaining electrodes are indicated in
  black. The electrodes from the remaining patients are colored by $k$-means
  cluster (computed using the full-brain correlation model shown in Panel D).
  \textbf{B.~Radial basis function kernel.} Each electrode contributed by the
  patient (black) weights on the full set of locations under consideration (all
  dots in Panel A, defined as $\bar{R}$ in the text).  The weights fall off with
  positional distance~\citep[in MNI152 space;][]{LyttEtal07} according to an RBF. \textbf{C.~Per-patient
  correlation matrices.}  After computing the pairwise correlations between the
  recordings from each patient's electrodes, we use RBF-weighted averages to
  estimate correlations between all locations in $\bar{R}$.  We obtain an
  estimated full-brain correlation matrix using each patient's data.
  \textbf{D.~Merged correlation model.}  We combine the per-patient correlation
  matrices (Panel C) to obtain a single full-brain correlation model that
  captures information contributed by every patient.  Here we have sorted the
  rows and columns to reflect $k$-means clustering labels~\citep[using
  $k$=7;][]{YeoEtal11}, whereby we grouped locations based on their correlations
  with the rest of the brain (i.e., rows of the matrix displayed in the panel).
  The boundaries denote the cluster groups.  The rows and columns of Panel C
  have been sorted using the Panel D-derived cluster labels.
  \textbf{E.~Reconstructing activity throughout the brain.}  Given the observed
  recordings from the given patient (shown in black; held-out recording is shown
  in blue), along with a full-brain correlation model (Panel D), we use
  Equation~\ref{eqn:reconstruction} to reconstruct the most probable activity at
  the held-out location (red).} \label{fig:methods}
\end{figure}

We first applied fourth order Butterworth notch filters to remove 60 Hz ($\pm$
.5 Hz) line noise from every recording (from every electrode). Next, we
downsampled the recordings (regardless of the original samplerate) to 250 Hz.
This downsampling step served to both normalize for differences in sampling
rates across patients and to ease the computational burden of our subsequent
analyses.  We then excluded any electrodes that showed putative epileptiform
activity. Specifically, we excluded from further analysis any electrode that
exhibited a maximum kurtosis of 10 or greater across all of that patient's
recording sessions.  We also excluded any patients with fewer than 2 electrodes
that passed this criteria, as the SuperEEG algorithm requires measuring
correlations between 2 or more electrodes from each patient.  For Dataset 1,
this yielded clean recordings from 4168 electrodes implanted throughout the
brains of 67 patients (Fig.~\ref{fig:methods}A, colored dots); for Dataset 2,
this yielded clean recordings from 5023 electrodes implanted in 78 patients.
Each individual patient contributed electrodes from a limited set of brain
locations, which we localized in a common space~\citep[MNI152;][]{GrabEtal06};
an example Dataset 1 patient's 54 electrodes that survived the kurtosis
thresholding procedure are highlighted in black and red
(Fig.~\ref{fig:methods}A).

@@

We designed our approach to be agnostic to electrode impedances, as electrodes
that do not exist do not have impedances.  Therefore our algorithm recovers
voltages in standard deviation ($z$-scored) units rather than attempting to
recover absolute voltages. (This property reflects the fact that
$\hat{K}_{\beta_s, \alpha_s}$ and $\hat{K}_{\alpha_s, \alpha_s}$ are correlation
matrices rather than covariance matrices.)  Also, we note that
Equation~\ref{eqn:reconstruction} requires computing a $T$ by $T$ matrix, which
can become computationally expensive when $T$ is very large (e.g., for the
Dataset 1 patient with the longest recording time, $T = 12,786,750$; also see
Fig.~\suppstats, Panels E and K). However, because
Equation~\ref{eqn:reconstruction} is time invariant, we may compute
$Y_{s,k,\beta_s}$ in a piecewise manner by filling in $Y_{s,k,\beta_s}$ one row
at a time (using the corresponding samples from $Y_{s, k, \alpha_s}$).

@@

We used this general approach to quantify the algorithm's performance across the
full dataset. For each held-out electrode, from each held-out patient in turn,
we computed the average correlation (across recording sessions) between the
SuperEEG-reconstructed voltage traces and the observed voltage traces from that
electrode.  For this analysis we set $\bar{R}$ to be the union of all electrode
locations across all patients.  This yielded a single correlation coefficient
for each electrode location in $\bar{R}$, reflecting how well the SuperEEG
algorithm was able to recover the recording at that location by incorporating
data across patients (black histogram in Fig.~\ref{fig:corrmap}A, map in
Fig.~\ref{fig:corrmap}C).  The observed distribution of correlations was
centered well above zero (mean: $r = 0.51$; $t$-test comparing mean of distribution of
$z$-transformed average patient correlation coefficients to 0: $t(66) = 23.55, p
< 10^{-10}$), indicating that the SuperEEG algorithm recovers held-out activity
patterns substantially better than random guessing.

Next, we compared the quality of these across-participant reconstructions (i.e.,
computed using a correlation model learned from other patients' data) to
reconstructions generated using a correlation model trained using the
in-patient's data.  In other words, for this within-patient benchmark analysis
we estimated $\hat{C}_{s}$ (Eqn.~\ref{eqn:subj_corrmat}) for each patient in
turn, using recordings from all of that patient's electrodes except at the
location we were reconstructing.  These within-patient reconstructions serve as
an estimate of how well data from all of the other electrodes from that single
patient may be used to estimate held-out data from the same patient.  This
allows us to ask how much information about the activity at a given electrode
might be inferred through (a) volume conductance or other sources of ``leakage''
from activity patterns measured from the patient's other electrodes and (b)
across-electrode correlations learned from that single patient.  As shown in
Figure~\ref{fig:corrmap}A (gray histogram), the distribution of within-patient
correlations was centered well above zero (mean: $r = 0.32$; $t$-test comparing
mean of distribution of $z$-transformed average patient correlation coefficients
to 0: $t(66) = 15.16, p < 10^{-10}$). However, the across-patient correlations
were substantially higher ($t$-test comparing average $z$-transformed within
versus across patient electrode correlations: $t(66) = 9.17, p < 10^{-10}$).
This is an especially conservative test, given that the across-patient SuperEEG
reconstructions exclude (from the correlation matrix estimates) all data from
the patient whose data is being reconstructed.  We repeated each of these
analyses on a second independent dataset and found similar results
(Fig.~\ref{fig:corrmap}B, D; within versus across reconstruction accuracy:
$t(77) = 11.25, p < 10^{-10}$). We also replicated this result separately for
each of the two experiments from Dataset 2 (Fig.~\perexpcorrmaps).  This overall
finding, that reconstructions of held-out data using correlation models learned
from \textit{other} patient's data yield higher reconstruction accuracy than
correlation models learned from the patient whose data is being reconstructed,
has two important implications.  First, it implies that distant electrodes
provide additional predictive power to the data reconstructions beyond the
information contained solely in nearby electrodes.  This follows from the fact
that each patient's grid, strip, and depth electrodes are implanted in a unique
set of locations, so for any given electrode the closest electrodes in the full
dataset tend to come from the same patient.  Second, it implies that the
spatial correlations learned using the SuperEEG algorithm are, to some extent,
similar across people.

@@

The recordings we analyzed from Dataset 1 comprised data collected as the
patients performed a variety of (largely idiosyncratic) tasks throughout each
day's recording session.  That we observed reliable reconstructions  across
patients suggests that the spatial correlations derived from the SuperEEG
algorithm are, to some extent, similar across tasks.  We tested this finding
more directly using Dataset 2.  In Dataset 2, the recordings were limited to
times when each patient was participating in one of two experiments.  Experiment
1 is a random-word list free recall task; Experiment 2 is a categorized list
free recall task (24 patients particpated in both).  We wondered whether a
correlation model learned from data from one experiment might yield good
predictions of data from the other experiment.  Further, we wondered about the
extent to which it might be beneficial or harmful to combine data across tasks.

To test the task-specificity of the SuperEEG-derived correlation models, we
restricted the dataset to the 24 patients that participated in both experiments
and repeated the above within- and across-patient cross validation procedures
separately for Experiment 1 and Experiment 2 data from Dataset 2.  We then
compared the reconstruction accuracies for held-out electrodes, for models
trained within versus across the two experiments, or combining across both
experiments (Fig.~\perexptaskrecon).  In every case we found that across-patient
models trained using data from all other patients out-performed within-patient
models trained on data only from the subject contributing the given electrode
($t$s$(23) > 6.50, p$s$ < 10^{-5}$). All reconstruction accuracies also reliably
exceeded chance performance ($t$s$(23) > 8.00, p$s$ < 10^{-8}$).  Average
reconstruction accuracy was highest for the across-patient models limited to
data from the same experiment (mean accuracy: $r = 0.68$); next-highest for the
models that combined data across both experiments (mean accuracy: $r = 0.61$); and
lowest for models trained across tasks (mean accuracy: $r = 0.47$).  This pattern of
results also held for each of the Dataset 2 experiments individually
(Fig.~\perexptaskreconseparated).  Taken together, these results indicate that
there are reliable commonalities in the spatial correlations of full-brain
activity across tasks, but that there are also reliable differences in these
spatial correlations across tasks. Whereas reconstruction accuracy benefits from
incorporating data from other patients, reconstruction accuracy is highest when
constrained to within-task data, or data that includes a variety of tasks (e.g.,
Dataset 1, or combining across the two Dataset 2 experiments).

\begin{figure}
  \centering \includegraphics[width=\textwidth]{figs/density}
  \caption{\textbf{Electrode sampling density by location.} \textbf{A.~Electrode
  sampling density by voxel location in Dataset 1.} Each voxel is colored by the
  proportion of total electrodes in the dataset that are located within 20~MNI
  units of the given voxel.  \textbf{B.~Electrode sampling density by voxel
  location in Dataset 2.}  This panel displays the sampling density map for
  Dataset 2, in the same format as Panel A. \textbf{C.~Correspondence in
  sampling density by voxel location across Datasets 1 and 2.}  The
  two-dimensional histogram displays the per-voxel sampling densities in the two
  Datasets, and the one-dimensional histograms display the proportions of voxels
  in each dataset with the given density value.  The correlation reported in the
  panel is across voxels in the 4~mm$^3$ MNI152 brain.} \label{fig:density}
\end{figure}

Although both datasets we examined provide good full-brain coverage (when
considering data from every patient), electrodes were not sampled uniformly
throughout the brain.  For example, in our patient population, electrodes are
more likely to be implanted in regions like the medial temporal lobe (MTL), and
are rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel MNI152 brain,
we computed the proportion of electrodes in the dataset that were contained
within a 20 MNI unit radius sphere centered on that voxel.  We defined the
\textit{density} at that location as this proportion. Across Datasets 1 and 2,
the electrode placement densities were similar (correlation by voxel: $r = 0.6,
p < 10^{-10}$).  We wondered whether regions with good covererage might be
associated with better reconstruction accuracy. For example,
Figures~\ref{fig:corrmap}A and B indicate that many electrodes in the MTL (which
tends to be relatively densely sampled) have relatively high reconstruction
accuracy, and occipital electrodes (which tends to be relatively sparsely
sampled) tend to have relatively low reconstruction accuracy.  To test whether
this held more generally across the entire brain, for each dataset we computed
the electrode placement density for each electrode from each patient (using the
proportion of \textit{other} patients' electrodes within 20 MNI units of the
given electrode).  We then correlated these density values with the
across-patient reconstruction accuracies for each electrode.  We found no
reliable correlation between reconstruction accuracy and density for Dataset 1
($r = 0.05, p = 0.70$) and a reliable negative correlation for Dataset 2 ($r =
-0.21, p = 0.05$).  This indicates that the reconstruction accuracies we
observed are \textit{not} driven solely by sampling density, but rather may also
reflect higher order properties of neural dynamics such as functional
correlations between distant voxels~\citep{BetzEtal17b}.

Prior work in humans and animals has shown that the spatial profile of the local
field potential differs by frequency band~\citep[e.g., with respect to volume
conductance properties and contribution to the local field
potential;][]{BuzsEtal12, FrieEtal07, CronEtal11}.  For example, lower frequency
components of the local field potential tend to have higher power and extend
further in space than high-frequency components~\citep[e.g., ][]{MillEtal07,
MannEtal09}.  We wondered whether the reconstructions we observed might be
differently weighting or considering the contributions of activity at different
frequency bands.  We therefore examined a range of frequency bands ($\delta$:
2--4 Hz; $\theta$: 4--8 Hz; $\alpha$: 8--12 Hz; $\beta$: 12--30 Hz; $\gamma_L$:
30--60 Hz; and $\gamma_H$: 60--100 Hz), along with a measure of broadband power.
We used second-order Butterworth bandpass filters to compute the activity
patterns within each narrow frequency band.  We defined broadband power as the
mean height of a linear robust regression fit in log-log space to the order 4
Morelet wavelet-computed power spectrum at 50 log-spaced frequencies from from
2--100 Hz~\citep{MannEtal09}. We then repeated our within-subject and
across-subject cross-validated reconstruction accuracy tests (analogous to
Fig.~\ref{fig:corrmap}) separately for each frequency band
(Fig.~\ref{fig:frequency}). (We also carried out a similar analysis on the
Hilbert transform-computed spectral power within each narrow band; see
Fig.~\freqpower.) Across both datasets, we found that our approach is best at
reconstructing patterns of broadband activity (right-most bars in
Figs.~\ref{fig:frequency}A and D), a correlate of population firing
rate~\citep{MannEtal09}.  We also achieved good reconstruction accuracy within
each narrow frequency band (Figs.~\ref{fig:frequency} and \freqpower).  Activity
at lower frequencies ($\delta$, $\theta$, $\alpha$, and $\beta$) tended to be
reconstructed better than high-frequency patterns ($\gamma_L$ and $\gamma_H$),
with reconstruction accuracy peaking in the $\alpha$ band.  Overall, these
results indicate that our approach is able to accurately recover information
within the 2--100 Hz range.
